---
title: "discussion_2_key"
format: docx
---

## Question 1:

You are studying a population of flowers that have 3 alleles at the flower color locus: A1, A2, and A3. You sample 1000 plants and genotype them at the locus. The genotype counts are: 290 A1A1, 100 A2A2, 70 A3A3, 410 A1A2, 100 A1A3, 230 A2A3. Is this flower color locus in Hardy-Weinberg equilibrium?


## Question 2:

You are studying the evolution of a lethal recessive allele in a population of anoles. The homozygous recessive genotype is completely lethal, the heterozygote and dominant homozygote are fine. You find that the allele has a frequency in the population of 0.01.

### 2.1

What is the ratio of heterozygotes to homozygotes in this population?

$$
2pq:q^2 = 198:1
$$

### 2.2 

How do you expect selection to change this ratio? What does this mean for selections ability to remove deleterious recessives?

**As selection removes the deleterious allele, relatively more of them will occur in heterozygotes than homozygotes meaning that they are invisible to selection. So selection cannot remove a deleterious recessive allele. **


It is often useful to code genetic data numerically. We often do this by counting the number of alleles in a genotype. For example, if we are studying a locus with two alleles---A1 and A2---we can count the number of A2 alleles. In the A1A1 genotype, there are no A2 alleles so it is coded as 0, likewise A1A2 and A2A2 are coded as 1 and 2 respectively. For the remaining questions, genotypes will be coded in this manner in the data (i.e. representing the count of A2 alleles)



## Question 3:

Analyze the data in the question 3 sheet of the Google sheet. The sheet represents an experiment where we transplanted 100 monkeyflowers of known genotype at a flower color locus. We came back and measured whether or not they survived (coded as 1 for survive and 0 for died).

```{r}
#| include: false
library(tidyverse)

w_bar <- function(p, w11, w12, w22){
  q <- 1-p
  p^2*w11 + 2*p*q*w12 + q^2*w22
}

p_prime <- function(p, w11, w12, w22){
  q <- 1-p
  (p/w_bar(p, w11, w12, w22))*(p*w11 + q*w12)
}

d <- read.csv("q3.csv")

```


### 3.1 What is the change in the frequency of the A1 allele if it's starting frequency is 0.8?

```{r}
#| echo: false
d %>% 
  group_by(genotype) %>% 
  summarise(survival = mean(survive)) %>% 
  mutate(relative_fit = round(survival/max(survival),3))
```

```{r}
#| echo: false
p_prime(.8, .778, .994, 1) - .8
```


### 3.2 What is the change in the frequency of the A1 allele if it's starting frequency is 0.2?

```{r}
#| echo: false
p_prime(.2, .778, .994, 1) - .2
```


### 3.3 Draw the adaptive topography for this fitness regime.


```{r}
#| echo: false
curve(w_bar(x, .778, .994, 1), 0, 1, xlab = "p", ylab = "Mean Fitness",
      lwd = 2)
points(c(1, w_bar(.5, .778, .994, 1), .778) ~ c(0, .5, 1), pch = 20, cex = 2)
```


## Question 4:

We have done the same experiment as question 3, but measured survival in a new environment. The data is under the question 4 tab in the Google sheet.

### 4.1 What is the change in the frequency of the A1 allele if it's starting frequency is 0.8?

```{r}
#| echo: false
d <- read.csv("q4.csv")

d %>% 
  group_by(genotype) %>% 
  summarise(survival = mean(survive)) %>% 
  mutate(relative_fit = round(survival/max(survival),3))
```

```{r}
#| echo: false
p_prime(.8, .6, 1, .844) - .8
```

### 4.2 What is the change in the frequency of the A1 allele if it's starting frequency is 0.2?

```{r}
#| echo: false
p_prime(.2, .6, 1, .844) - .2
```



### 4.3 Draw the adaptive topography for this fitness regime.

**First calculate** $\hat{p} = \frac{w_{22} - w_{12}}{w_{11}+w_{22}-2w_12}$

```{r}
(p_hat = (.844 - 1)/(.844 + .6 - 2))
```

```{r}
#| echo: false
curve(w_bar(x, .6, 1, .844), 0, 1, xlab = "p", ylab = "Mean Fitness",lwd = 2)
points(c(.844, w_bar(p_hat, .6,1,.844), .6) ~ c(0, p_hat,1),
       pch = 20, cex = 2)
```


## Question 5:

We have done the same experiment as questions 3 and 4, but measured survival in a yet a new environment. The data is under the question 5 tab in the Google sheet.

### 5.1 What is the change in the frequency of the A1 allele if it's starting frequency is 0.8?


```{r}
#| echo: false
d <- read.csv("q5.csv")

d %>% 
  group_by(genotype) %>% 
  summarise(survival = mean(survive)) %>% 
  mutate(relative_fit = round(survival/max(survival),3))
```


```{r}
#| echo: false
p_prime(.8, 1, .565, .804) - .8
```




### 5.2 What is the change in the frequency of the A1 allele if it's starting frequency is 0.2?


```{r}
#| echo: false
p_prime(.2, 1, .565, .804) - .2
```
### 5.3 Draw the adaptive topography for this fitness regime.

**Calculate** $\hat{p}$.

```{r}
(p_hat = (.804 - .565)/(1 + .804 - 2*.565))
```

```{r}
#| echo: false
curve(w_bar(x, 1, .565, .804), xlab = "p", ylab = "Mean Fitness",lwd = 2)
points(c(.804, w_bar(p_hat, 1,.565,.804), 1) ~ c(0, p_hat,1),
       pch = 20, cex = 2)
```






